# Interaction

## Example 1

```{r}
df=data.frame(y=c(3, 3, 3, 7, 8, 5, 6, 3, 5, 7), 
              trt=as.factor(c("A", "A", "A", "A", "A", "B", "B", "B", "B", "B")), 
              center=as.factor(c(1, 1,1, 2, 2, 1, 1, 2,2, 2)))


df
```

**The following three models are same.** 

```{r}
fit1=lm(y~trt*center, data=df)
fit2=lm(y~trt+center+trt*center, data=df)
fit3=lm(y~trt+center+trt:center, data=df)

```

```{r}
summary(fit1)

summary(fit2)

summary(fit3)

```

```{r}
model.matrix(fit1)

model.matrix(fit2)

model.matrix(fit3)
```

Model: 

\[y_i=\beta_0+\beta_1 I(i\in \text{trtB})+\beta_2 I(i\in \text{center2})+\beta_3 I(i \in \text{trtB and center2})+\varepsilon_i \]

- trtA & center2: $\beta_0+\beta_2$

- trtB & center2: $\beta_0+\beta_1+\beta_2+\beta_3$

- difference of trt effects conditioning on center2: $\beta_1+\beta_3$

- trtA & center1: $\beta_0$

- trtB & center1: $\beta_0+\beta_1$

- difference of trt effects conditioning on center1: $\beta_1$

- Therefore, there is an interaction between trt and center



## Example 2


```{r}
set.seed(12)
f1 <- gl(n = 2, k = 30, labels = c("Low", "High")) ## generate factor levels 
f2 <- as.factor(rep(c("A", "B", "C"), times = 20))
modmat <- model.matrix(~f1 * f2, data.frame(f1 = f1, f2 = f2)) ## make the design matrix
coeff <- c(1, 3, -2, -4, 1, -1.2)
y <- rnorm(n = 60, mean = modmat %*% coeff, sd = 0.1)
dat <- data.frame(y = y, f1 = f1, f2 = f2)


summary(lm(y ~ f1 * f2))
model.matrix(lm(y ~ f1 * f2))

```

Model: 
\[y_i=\beta_0+\beta_1I(i\in \text{f1High})+\beta_2I(i\in \text{f2B})+\beta_3I(I\in \text{f2C})+ \\
\beta_4 I(i\in \text{f1High and f2B})+\beta_5 I(i\in\text{f1High and f2C})+\varepsilon_i \]

- f1 high and f2 A: $\beta_0+\beta_1$

- f1 high and f2 B: $\beta_0+\beta_1+\beta_2+\beta_4$

- f1 high and f2 C: $\beta_0+\beta_1+\beta_3+\beta_5$

- f1 low and f2 A: $\beta_0$

- f1 low and f2 B: $\beta_0+\beta_2$

- f1 low and f2 C: $\beta_0+\beta_3$

- See more at https://rpubs.com/hughes/15353
