# Least squares mean 

## Reference

- https://cran.r-project.org/web/packages/emmeans/vignettes/basics.html
- https://cran.r-project.org/web/packages/emmeans/vignettes/comparisons.html


## Pigs data example 

```{r}
library(emmeans)
library(tidyverse)
```

```{r}
with(pigs, interaction.plot(percent, source, conc))
pigs%>%group_by(percent)%>%summarise(marginalmean=mean(conc))
pigs%>%group_by(percent, source)%>%summarise(mean(conc))
```



```{r}
lm1=lm(conc ~ source + factor(percent), data = pigs)
summary(lm1)


```


```{r}
lsm=emmeans(lm1, specs = "source")
lsm
pairs(lsm)
```


Check how estimated marginal mean is calculated: 

```{r}
lm1coef=as.vector(coef(lm1))

# balanced design
lm1coef[1]+(lm1coef[4]+lm1coef[5]+lm1coef[6])/4
lm1coef[1]+lm1coef[2]+(lm1coef[4]+lm1coef[5]+lm1coef[6])/4
lm1coef[1]+lm1coef[3]+(lm1coef[4]+lm1coef[5]+lm1coef[6])/4
```


Add one more column in pigs dataset

```{r}
set.seed(123)
pigs$normalrv=rnorm(nrow(pigs), mean = 10, sd = 1)

lm2=lm(conc ~ source + factor(percent) + normalrv, data = pigs)
summary(lm2)
emmeans(lm2, specs = "source")
```


- Model: 

\[y_i=\alpha_0+\alpha_1 I(S_i=\text{soy})+\alpha_2 I(S_i=\text{skim})+\alpha_3 I(F_i=12)+\alpha_4 I(F_i=15)+ \\
\alpha_5 I(F_i=18)+ \alpha_6 X_i \]

- Assume a **balanced** design, we have 
\begin{equation*}
\begin{split}
E(y_i \mid S_i=\text{fish})=E_{X}\left\{\frac{1}{4}\left[E(y_i \mid S_i=\text{fish}, F_i=9,X)+E(y_i \mid S_i=\text{fish}, F_i=12,X)+ \\
E(y_i \mid S_i=\text{fish}, F_i= 15,X)+E(y_i \mid S_i=\text{fish}, F_i=18,X)\right]\right\}
\end{split}
\end{equation*}


```{r}
alpha0=as.vector(coef(lm2))[1]
alpha1=as.vector(coef(lm2))[2]
alpha2=as.vector(coef(lm2))[3]
alpha3=as.vector(coef(lm2))[4]
alpha4=as.vector(coef(lm2))[5]
alpha5=as.vector(coef(lm2))[6]
alpha6=as.vector(coef(lm2))[7]
meanX=mean(pigs$normalrv)

alpha0+(alpha3+alpha4+alpha5)/4+alpha6*meanX

alpha0+alpha1+(alpha3+alpha4+alpha5)/4+alpha6*meanX

alpha0+alpha2+(alpha3+alpha4+alpha5)/4+alpha6*meanX

```


## Iris data example 

```{r}
# make virginica the reference group 
iris <- iris %>%
  mutate( Species = forcats::fct_relevel(Species, 'virginica') )

m1 <- lm( Sepal.Width ~ Sepal.Length + Species, data=iris )  # Parallel Lines
summary(m1)

```

```{r}
emmeans(m1, specs = "Species")
```

**Calculate the estimated marginal mean as follows:**

- Model: $y_i=\alpha_0+\alpha_1X_i+\alpha_2 I(S_i=\text{setosa})+\alpha_3 I(S_i=\text{versicolor})$

- $E(y_i \mid S_i=\text{virginica})=E_{X}(E(y_i \mid S_i=\text{virginica}, X))=\alpha_0+\alpha_1 E(X)$
- $E(y_i \mid S_i=\text{setosa})=E_{X}(E(y_i \mid S_i=\text{setosa}, X))=\alpha_0+\alpha_1 E(X)+\alpha_2$
- $E(y_i \mid S_i=\text{versicolor})=E_{X}(E(y_i \mid S_i=\text{versicolor}, X))=\alpha_0+\alpha_1E(X)+\alpha_3$


```{r}
alpha0=as.vector(coef(m1))[1]
alpha1=as.vector(coef(m1))[2]
alpha2=as.vector(coef(m1))[3]
alpha3=as.vector(coef(m1))[4]
meanX=mean(iris$Sepal.Length)

alpha0+alpha1*meanX
alpha0+alpha1*meanX+alpha2
alpha0+alpha1*meanX+alpha3

```




